---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, include = F}
library(survival)
library(rjags)
library(SHELF)
```


# DTEAssurance

<!-- badges: start -->
<!-- badges: end -->

DTEAssurance is an R package  for implementing the assurance method for a delayed treatment effect. Distributions are elicited for the length of delay and the post-delay hazard ratio. The package implements the methodology from  


* [Salsbury JA, Oakley JE, Julious SA, Hampson LV. Assurance methods for designing a clinical trial with a delayed treatment effect. Statistics in Medicine. 2024; 43(19): 3595-3612. doi: 10.1002/sim.10136](https://onlinelibrary.wiley.com/doi/10.1002/sim.10136)


## Installation

You can install the development version of DTEAssurance from [GitHub](https://github.com/) with:

```{r}
devtools::install_github("jamesalsbury/DTEAssurance")
```

```{r}
library(DTEAssurance)
```

## Example (from the paper)

### Control

```{r}
ZODIAC <- read.csv("data/ZODIAC.csv")
REVEL <- read.csv("data/REVEL.csv")
INTEREST <- read.csv("data/INTEREST.csv")
```


```{r}

ZodiacFit <- survfit(Surv(Survival.time, Status)~1, data = ZODIAC)
plot(ZodiacFit, col = "blue", conf.int = F, xlab = "Time (months)", ylab = "Overall Survival")

RevelFit <- survfit(Surv(Survival.time, Status)~1, data = REVEL)
lines(RevelFit, col = "red", conf.int = F)

InterestFit <- survfit(Surv(Survival.time, Status)~1, data = INTEREST)
lines(InterestFit, col = "yellow", conf.int = F)

combinedControlArms <- rbind(INTEREST, REVEL, ZODIAC)

combinedFit <- survfit(Surv(Survival.time, Status)~1, data = combinedControlArms)
lines(combinedFit, col = "black", conf.int = F)

legend("topright", legend = c("ZODIAC", "REVEL", "INTEREST", "Combined"), lty = 1, col = c("blue", "red", "yellow", "black"))

```


```{r}
#Performing MCMC on this data set

modelstring="

data {
  for (j in 1:n){
    zeros[j] <- 0
  }
}

model {
  C <- 10000
  for (i in 1:n){
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, log(gamma_C)+gamma_C*log(lambda_C*datTimes[i])-(lambda_C*datTimes[i])^gamma_C-log(datTimes[i]), -(lambda_C*datTimes[i])^gamma_C)
  }
  
    lambda_C ~ dnorm(1,1/10000)T(0,)
    gamma_C ~ dnorm(1,1/10000)T(0,)
    
    }
"

model = jags.model(textConnection(modelstring), data = list(datTimes = combinedControlArms$Survival.time, datEvent = combinedControlArms$Status, n= nrow(combinedControlArms)), quiet = T) 

update(model, n.iter=1000)
output=coda.samples(model=model, variable.names=c("lambda_C", "gamma_C"), n.iter = 10000)

plot(output)

lambda_Csample <- as.numeric(unlist(output[,2]))
gamma_Csample <- as.numeric(unlist(output[,1]))
```

### Eliciting T

First, we use SHELF to elicit beliefs about T, the length of the delay

```{r}
P_S <- 0.9
P_DTE <- 0.7
TBeliefs <- SHELF::fitdist(c(3, 4, 5), probs = c(0.25, 0.5, 0.75), lower = 0, upper = 10)
```


### Eliciting post-delay HR

and the treatment effect, in the form of the post-delay hazard ratio

```{r}
HRStarBeliefs <- SHELF::fitdist(c(0.55, 0.6, 0.7), probs = c(0.25, 0.5, 0.75), lower = 0, upper = 1.5)
```





