---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DTEAssurance

<!-- badges: start -->
<!-- badges: end -->

DTE is an R package which elicits two distributions and then calculates assurance using these elicited prior distributions. The package implements the methods as described in 

* [this paper](https://jamesalsbury.github.io/)


## Installation

You can install the development version of DTEAssurance from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jamesalsbury/DTEAssurance")
```

## shiny app

The only function contained within the package is one which launches a `shiny` app. The app allows you to implement all of the methods outlined in [this paper](https://jamesalsbury.github.io/). To launch the app, run

```{r, eval = F}
DTEAssurance::DTEAssuranceApp()
```
